# PFR Analysis {#sec-4_pfr_analysis}

## PFR Characteristics

## PFR Operation

## Qualitative Analysis of Steady-State Reaction in a PFR

## PFR Design Equations

## PFR Reaction Fronts

## Packed Bed PFRs

## Examples

```{r}
#| echo: false
#| output: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readxl)
source("~/Libraries/R/fmt_tibble_col.R")
```


### Steady-State Conversion in a Gas-Phase PFR

{{< include examples/reb_13_7_1/narrative.qmd >}}

{{< include examples/reb_13_7_1/equations.qmd >}}

---

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point
identify problem type, describe variable naming
:::

#### Assignment Summary

**Given and Known Constants**: $\Delta H_1$ = 44.8 kJ mol^-1^, $k_{0,1}$ = 7.22 x 10^6^ mol atm^-2^ cm^-3^ s^-1^, $E_1$ = 84.1 kJ mol^-1^, $L$ = 10 ft, $D$ = 1 in, $T_{ex}$ = 200 °C, $U$ = 7.48 x 10^4^ J h^-1^ ft^-2^ K^-1^, $y_{A,in}$ = 0.60, $y_{B,in}$ =  0.40, $\dot{V}_{in}$ = 282 L min^-1^, $P$ = 2.5 atm, $T_{in}$ = 175 °C, $\hat{C}_{p,A}$ = 18.0 cal mol^-1^ K^-1^, $\hat{C}_{p,B}$ = 12.25 cal mol^-1^ K^-1^, $\hat{C}_{p,Z}$ = 21.2 cal mol^-1^ K^-1^, and $R$ = 8.314 J mol^-1^ K^-1^ = 0.08206 L atm mol^-1^ K^-1^.

**Reactor System**: Heated, steady-state PFR

**Quantities of Interest**: $T \big\vert_{z=L}$ and $f_B$

#### Mathematical Formulation of the Solution

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point

@eq-pfr_ss_mol_balance

$$
\frac{d \dot{n}_i}{d z}  =\frac{\pi D^2}{4}\sum_j \nu_{i,j}r_j
$$

@eq-pfr_ss_energy_balance

$$
\left(\sum_i \dot{n}_i \hat{C}_{p,i} \right) \frac{d T}{d z} = \pi D U\left( T_{ex} - T  \right) - \frac{\pi D^2}{4}\sum_j r_j \Delta H_j
$$

:::

**Reactor Design Equations**

Mole balance design equations for A, B, and Z are presented in equations (3) - (5). The energy balance on the reacting gas is given in equation (6).

$$
\frac{d \dot{n}_A}{dz} = -\frac{\pi D^2}{4}r_1 \tag{3}
$$

$$
\frac{d \dot{n}_B}{dz} = -\frac{\pi D^2}{4}r_1 \tag{4}
$$

$$
\frac{d \dot{n}_Z}{dz} = \frac{\pi D^2}{4}r_1 \tag{5}
$$

$$
\frac{d T}{d z} = \frac{\pi D U\left( T_{ex} - T  \right) - \frac{\pi D^2}{4}r_1 \Delta H_1}{\left( \dot{n}_A \hat{C}_{p,A} + \dot{n}_B \hat{C}_{p,B} + \dot{n}_Z \hat{C}_{p,Z} \right)} \tag{6}
$$

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point
check number of ivodes and deps
:::

differential ideal gas law

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point
define initial values and stopping criterion
:::

**Initial Values and Stopping Criterion**

| Variable | Initial Value | Stopping Criterion |
|:------:|:-------:|:-------:|
| $z$ | $0$ | $L$ |
| $\dot{n}_A$ | $\dot{n}_{A,in}$ | |
| $\dot{n}_B$ | $\dot{n}_{B,in}$ | |
| $\dot{n}_Z$ | $0$ | |
| $T$ | $T_{in}$ | |
  
: Initial values and stopping criterion for solving  equations, (numbers). {#tbl-example__initial_values}

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point

In order to solve the IVODEs numerically I'll need to do two things: calculate the values of the derivatives at the start of each integration step and calculate all of the initial and final values in @tbl-example__initial_values.

I'll start by writing the design equations in the form of derivative expressions (see @sec-apndx....

I'll need to evaluate the derivatives at the start of each integration step. At that point, the independent and dependent variables will be known along with the given and known constants. I will need to calculate any other quantities that appear in the derivatives expressions.
:::

**Ancillary Equations for Evaluating the Derivatives**

$$
k_1 = k_{0,1} \exp{\left( \frac{-E_1}{RT} \right)} \tag{7}
$$

$$
P_A = \frac{\dot{n}_A}{\dot{n}_A + \dot{n}_B + \dot{n}_Z}P \tag{8}
$$

$$
P_B = \frac{\dot{n}_B}{\dot{n}_A + \dot{n}_B + \dot{n}_Z}P \tag{9}
$$

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point
When I solve the IVODEs numerically, I'll also need to know the initial and final values in @tbl-example__initial_values....
:::

**Ancillary Equations for Calculating the Initial and Final Values**

$$
\dot{n}_{A,in} = y_{A,in}\frac{P\dot{V}_{in}}{RT_{in}}\tag{10}
$$

$$
\dot{n}_{B,in} = y_{B,in}\frac{P\dot{V}_{in}}{RT_{in}}\tag{11}
$$

:::{.callout-tip collapse="true"}
## Click Here to See What an Expert Might be Thinking at this Point
develop equations needed to calculate quantities of interest
:::

**Ancillary Equations for Calculating the Quantities of Interest**

Solving the reactor design equations yields [list independent and dependent variables]. 

$$
f_B = \frac{\dot{n}_{B,in} - \dot{n}_B}{\dot{n}_{B,in}}\tag{12}
$$

**Calculations Summary**

1. Substitute given and known constants into all equations.
2. When it is necessary to evaluate the derivatives
	a. [list of variables] will be available.
	b. 
3. When it is necessary to calculate the initial and final values in @tbl-example__initial_values
	a. [list of variables] are known constants.
	b. 
4. When it is necessary to evaluate the residual corresponding to [variable]
    a. a guess for [variable] will be available for use in all equations.
    b. 
5. When it is necessary to calculate the quantities of interest, [list]
	a. corresponding sets of values of [independent variable], [dependent variable], spanning the range from their initial values to their final values will be available.
	b. 

#### Numerical implementation of the Solution

1. Make the given and known constants available for use in all functions.
2. Define a variable to hold [unknown constant] and make it available to all functions.
3. Write a derivatives function that
	a. receives the independent and dependent variables, [list], as arguments,
	b. evaluates the derivatives as described in step 2 of the calculations summary, and 
	c. returns the values of the derivatives, [list].
4. Write a residuals function that
	a. receives a guess for [variable] and makes it available to all functions,
	b. evaluates the residual as described in step 4 of the calculations summary, and
	c. returns the value of [variable].
5. Write a reactor model function that
	a. ...
	b. calculates the initial and final values in @tbl-example__initial_values as described in step 3 of the calculations summary,
	c. gets corresponding sets of values of [independent variable], [dependent variable], spanning the range from their initial values to their final values by calling an IVODE solver and passing the following information to it
		i. the initial values and stopping criterion in @tbl-example__initial_values and
		ii. the name of the derivatives function from step 3 above,
	d, checks that the solver successfully solved the IVODEs, and
	e. returns the values returned by the IVODE solver.
6. Perform the analysis by
	a. ...
	b. getting corresponding sets of values of [independent variable], [dependent variable], spanning the range from their initial values to their final values by calling the reactor model function from step 5 above,
	c. calculates the quantities of interest as described in step 5 of the calculations summary,
	d. ....

#### Results and Discussion

```{r}
#| echo: false
#| output: false
df <- read.csv("examples/reb_13_7_1/results.csv") 
df <- fmt_tibble_col(df, 2, 3, 3, 1)
```

The calculations were performed as described above. The outlet temperature is `r df$value[1]` `r df$units[1]` and the conversion of B is `r df$value[2]` `r df$units[2]`.


::: {.callout-note appearance="simple"}
## [SC]{style="color:blue"}o[RE]{style="color:red"} Connection
Videos showing how to complete this assignment using either Matlab or Python, along with the Matlab and Python code, are available in [SCoRE](http://buffalobadger.github.io/SCoRE/class_19.html).
:::

## Symbols Used in [Chapter -@sec-4_pfr_analysis]
